Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants

### 2016-10-14 paper/paper.tex
Fig. -> Figure for uniformity

By applying a state-of-the-art CEGIS engine we are now able to present a
tool that automatically \emph{generates} digital controllers for a given
continuous plant model that are correct by construction.  This approach promises to
reduce the cost and time of development of digital control dramatically, and
requires considerably less expertise than methodologies driven by
verification.  Specifically, we synthesise controllers for hybrid
closed-loop systems, i.e., software-implemented embedded controllers along
with a model of their physical environment (the plant), that are
\emph{stable}.  Due to the complexity of such systems, in this work we focus on linear
models with known configurations and perform parametric synthesis of stabilising digital controllers.

Our work addresses particularly challenging aspects of the control synthesis
problem.  We perform digital control synthesis over a hybrid model, where
the plant exhibits continuous behaviour whereas the controller operates in 
discrete time and over a quantised domain.  
Inspired by a classical approach \cite{astrom1997computer}, 
we translate the problem into a single digital domain.  
Given our primary interest in stability despite digitisation effects, we have modelled
a digital equivalent of the plant by evaluating the effects of the
quantisers (A/D and D/A converters) as time discretisation elements with
quantisation noise.  The resulting closed-loop system is a program loop that
operates on bit-vectors encoded using fixed-point arithmetic with fixed word
length (FWL). The deterministic effects of the FWL and
quantisation errors are incorporated into the model as an additive
uncertainty, which is taken into account during the CEGIS-based synthesis of
the control software for the plant.

@@

\item We design and implement an automatic approach for generating
  {\em corect-by-construction} digital controllers that is based on
  state-of-the-art inductive synthesis techniques.  Our application of
  program synthesis is non-trivial and requires addressing challenges
  specific to embedded control systems, such as the effects of quantisers and
  FWL. In particular, we have found that a two-stage verification engine
  that continuously refines the precision of the fixed-point
  representation is key to the performance of our approach yielding,
  in most cases, to a speed-up of at least two orders of magnitude over a
  conventional one-stage verification engine.

@@

The digital controllers synthesised using the algorithm we present in this
paper are typically used in closed loops with continuous (physical)
plants.  Thus, the proposed approach has to consider dynamics of continuous
parts of the system (i.e., the plant) and the discrete parts (i.e., the
digital controller).  To obtain a joint model, we fully discretise the continuous
plant. We briefly recall the standard foundations of this step.

@@

\begin{mylemma}{\cite{Bessa16,fadali}}
\label{lemma:eq_int_stab} 
%
A feedback hybrid control system as given in Fig.~\ref{fig:sampledsystem},
with a FWL implementation of the digital controller $\hat{C}(z) =
\frac{\mathit{FWL}[N_{C}(z)]}{\mathit{FWL}[D_{C}(z)]}$ and uncertain
discrete model of the plant $\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q
\vec{G} \Rightarrow \hat{G}(z)=\frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$, is
BIBO-stable if and only if:
%
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
S(z)=\mathit{FWL}[N_{C}(z)]\hat{N}_{G}(z)+\mathit{FWL}[D_{C}(z)]\hat{D}_{G}(z);
\end{equation}
\item the direct loop product $\frac{\mathit{FWL}[N_{C}(z)]}{\mathit{FWL}[D_{C}(z)]}\cdot \frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{mylemma}

@@

In order to synthesise closed-loop digital control systems, we use a program
synthesis engine.  Our program synthesiser implements Counter-Example Guided
Inductive Synthesis (CEGIS)~\cite{sketch}.  We start by presenting its
general architecture followed by describing the parts specific to
closed-loop control systems.  A high-level view of the synthesis process is
given in Fig.~\ref{DSSynth_process}.  Steps $1$ to $3$ are performed by
the user and Steps~A to D are automatically performed by our tool for
Digital Systems Synthsis, named DSSynth.
%
%, which accepts the
%digital controller transfer functions together with the plant model.  For
%any digital controller, implementation details should be provided ({\it
%e.g.}, number of bits, realization, and sample time).

@@

The design of our synthesiser is given in Fig.~\ref{fig:CEGIS} and consists
of two phases, {\sc Synthesise} and {\sc Verify}, which interact via a
finite set of test vectors {\sc inputs} that is updated incrementally. 
Given the aforementioned specification $\sigma$, the {\sc synth} procedure
tries to find an existential witness $\vec{P}$ satisfying the specification
$\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to
all $\vec{x} \in \mathcal{D}$).
%
If {\sc synthesise} succeeds in finding an witness $\vec{P}$, this witness
is a candidate solution to the full synthesis formula.  We pass this
candidate solution to {\sc verify}, which checks whether it is a full
solution ({\it i.e.}, $\vec{P}$ satisfies the specification $\sigma(\vec{x},
\vec{P})$ for all $\vec{x}\in\mathcal{D}$).
%
%determines whether
%it does satisfy the specification on all inputs by checking
%satisfiability of the following verification formula:
%
%$\exists \vec{x} . \lnot \sigma(\vec{x}, \vec{P})$
%If this formula is unsatisfiable, 
%% the candidate solution is in fact a 
%% full solution.
%
If this is the case, then the algorithm terminates.  Otherwise, additional
information is provided to the {\sc synthesise} phase in the form of a new
counterexample that is added to the {\sc inputs} set and the loop iterates
again (note that, for now, we ignore the second feedback signal ``Increase
Precision'' provided by the {\sc Verify} phase in Fig.~\ref{fig:CEGIS} as it
is specific to control synthesis and will be described in the next section).

@@

The second approach is illustrated in Fig.~\ref{fig:CEGIS} and uses a
two-stage verification: the first stage performs potentially unsound
fixed-point operations assuming a precision \param{I_b}{F_b}, and the
second stage restores soundness by validating these operations
using interval arithmetic on the synthesised controller.

In more detail, in the first stage, denoted by {\sc Uncertainty} in Fig.~\ref{fig:CEGIS}, 
assuming a precision \param{I_b}{F_b} 
we check whether the system is unstable for the current candidate solution,
{\it i.e.}, if $\neg \phi_{stability}$ is satisfiable for $S(z)$.
If this is the case, then we obtain a counterexample 
$\tilde G(z)$, 
%denoted by values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$, 
%\red{[see my comment above - these quantities pop up too abruptly.]},
which makes the closed-loop system unstable. 
This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $C(z)$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.

>>>>

By applying a state-of-the-art CEGIS engine we are now able to present a
tool that automatically \emph{generates} digital controllers for a given
plant model that are correct by construction.  This approach promises to
reduce the cost and time of development of digital control dramatically, and
requires considerably less expertise than methodologies driven by
verification.  Specifically, we synthesise controllers for hybrid
closed-loop systems, i.e., software-implemented embedded controllers along
with a model of their physical environment (the plant), that are
\emph{stable}.  Due to the complexity of such systems, we focus on linear
models with known configurations to perform parametric synthesis of the
controller dynamics.

Our work addresses particularly challenging aspects of the control synthesis
problem.  We perform digital control synthesis over a hybrid model, where
the plant exhibits continuous behaviour and the controller operates in both
discrete time and discrete state space.  The key insight of our approach is
that the problem can be translated into a \emph{single, joint domain}, which
represents a bi-simulation of the original, hybrid dynamics.  Given our
primary interest in stability despite digitisation effects, we have modelled
a digital equivalent of the plant by evaluating the effects of the
quantisers (A/D and D/A converters) as time discretisation elements with
quantisation noise.  The resulting closed-loop system is a program loop that
operates on bit-vectors encoded using fixed-point arithmetic with fixed word
length (FWL).  The deterministic effects of the FWL and
quantisation errors are incorporated into the model as an additive
uncertainty, which is taken into account during the CEGIS-based synthesis of
the control software for the plant.

@@

\item We designed and implemented an automatic approach for generating
  {\em corect-by-construction} digital controllers that is based on
  state-of-the-art inductive synthesis techniques.  Our application of
  program synthesis is non-trivial and requires addressing challenges
  specific to embedded control systems such as the effects of quantisers and
  FWL. In particular, we found that a two-stage verification engine
  that continuously refines the precision of the fixed-point
  representation is key to the performance of our approach yielding,
  in most cases, to a speed-up of at least two orders of magnitude over a
  conventional one-stage verification engine.

@@

The digital controllers synthesised using the algorithm we present in this
paper are typically used in closed loops with continuous (and physical)
plants.  Thus, the proposed approach has to consider dynamics of continuous
parts of the system (i.e., the plant) and the discrete parts (i.e., the
digital controller).  To obtain a joint model, we discretise the continuous
plant. We briefly recall the standard foundations of this step.

@@

\begin{mylemma}{\cite{Bessa16,fadali}}
\label{lemma:eq_int_stab} 
%
A feedback hybrid control system as given in Figure~\ref{fig:sampledsystem},
with a FWL implementation of the digital controller $\hat{C}(z) =
\frac{\mathit{FWL}[N_{C}(z)]}{\mathit{FWL}[D_{C}(z)]}$ and uncertain
discrete model of the plant $\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q
\vec{G} \Rightarrow \hat{G}(z)=\frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$, is
BIBO-stable if and only if:
%
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
S(z)=\mathit{FWL}[N_{C}(z)]\hat{N}_{G}(z)+\mathit{FWL}[D_{C}(z)]\hat{D}_{G}(z);
\end{equation}
\item the direct loop product $\frac{\mathit{FWL}[N_{C}(z)]}{\mathit{FWL}[D_{C}(z)]}\cdot \frac{\hat{N}_{G}(z)}{\hat{D}_{G}(z)}$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{mylemma}

@@

In order to synthesise closed-loop digital control systems, we use a program
synthesis engine.  Our program synthesiser implements Counter-Example Guided
Inductive Synthesis (CEGIS)~\cite{sketch}.  We start by presenting its
general architecture followed by describing the parts specific to
closed-loop control systems.  A high-level view of the synthesis process is
given in Figure~\ref{DSSynth_process}.  Steps $1$ to $3$ are performed by
the user and Steps~A to D are automatically performed by our tool for
Digital Systems Synthsis, named DSSynth.
%
%, which accepts the
%digital controller transfer functions together with the plant model.  For
%any digital controller, implementation details should be provided ({\it
%e.g.}, number of bits, realization, and sample time).

@@

The design of our synthesiser is given in Figure~\ref{fig:CEGIS} and consists
of two phases, {\sc Synthesise} and {\sc Verify}, which interact via a
finite set of test vectors {\sc inputs} that is updated incrementally. 
Given the aforementioned specification $\sigma$, the {\sc synth} procedure
tries to find an existential witness $\vec{P}$ satisfying the specification
$\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to
all $\vec{x} \in \mathcal{D}$).
%
If {\sc synthesise} succeeds in finding an witness $\vec{P}$, this witness
is a candidate solution to the full synthesis formula.  We pass this
candidate solution to {\sc verify}, which checks whether it is a full
solution ({\it i.e.}, $\vec{P}$ satisfies the specification $\sigma(\vec{x},
\vec{P})$ for all $\vec{x}\in\mathcal{D}$).
%
%determines whether
%it does satisfy the specification on all inputs by checking
%satisfiability of the following verification formula:
%
%$\exists \vec{x} . \lnot \sigma(\vec{x}, \vec{P})$
%If this formula is unsatisfiable, 
%% the candidate solution is in fact a 
%% full solution.
%
If this is the case, then the algorithm terminates.  Otherwise, additional
information is provided to the {\sc synthesise} phase in the form of a new
counterexample that is added to the {\sc inputs} set and the loop iterates
again (note that, for now, we ignore the second feedback signal ``Increase
Precision'' provided by the {\sc Verify} phase in Figure~\ref{fig:CEGIS} as it
is specific to control synthesis and will be described in the next section).

@@

The second approach is illustrated in Figure~\ref{fig:CEGIS} and uses a
two-stage verification: the first stage performs potentially unsound
fixed-point operations assuming a precision \param{I_b}{F_b}, and the
second stage restores soundness by validating these operations
using interval arithmetic on the synthesised controller.

In more detail, in the first stage, denoted by {\sc Uncertainty} in Figure~\ref{fig:CEGIS}, 
assuming a precision \param{I_b}{F_b} 
we check whether the system is unstable for the current candidate solution,
{\it i.e.}, if $\neg \phi_{stability}$ is satisfiable for $S(z)$.
If this is the case, then we obtain a counterexample 
$\tilde G(z)$, 
%denoted by values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$, 
%\red{[see my comment above - these quantities pop up too abruptly.]},
which makes the closed-loop system unstable. 
This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $C(z)$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.
